R Markdown Assignment - Peer Assessment 1
 - By Robert Adams
======================================

Let's first read the data into R using the directory path to where we saved the data. It is a CSV format so we use read.csv.

```{r}
path <- "C:/Users/Bobby/Google Drive/Data Science/Data Sets/activity.csv"
Data <- read.csv(file.path(path), header=T, sep=",", stringsAsFactors=FALSE)
```

Let's look at the head of the data to make sure it loaded in correctly

```{r}
head(Data)
```

Next we're going to omit the N/A data, as it is useless in this analysis

```{r}
Data <- na.omit(Data)
```

Now we set the date variable as a date class

```{r}
Data$date <- as.Date(Data$date)
```

Now we load the ggplot 2 package and plot the histogram for Number of Steps by date

```{r}
library(ggplot2)
g <- ggplot(Data, aes(Data$date, Data$steps))
g + geom_bar(stat="identity", colour = "steelblue", fill = "steelblue", width = 0.7) +
labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")
```

In order to report the mean steps by day we need to aggregate the data:

```{r}
Data1 <- aggregate(Data$steps~Data$date, FUN="mean")
```

and to report the median we do calculate the total number of steps for the day, then turn it into the median:

```{R}
Data2 <- aggregate(Data$steps~Data$date, FUN="sum")
median(Data2$`Data$steps`)
```

Now for a time series plot, averaged steps across all days in a time series plot for the 5 minute intervals.

```{r}
Data3 <- aggregate(Data$steps~Data$interval, FUN="mean")
g <- ggplot(Data3, aes(Data3$`Data$interval`, Data3$`Data$steps`))
g + geom_line() +labs(title = "Time Series Plot of the 5-minute Interval", x = "5-minute intervals", y = "Average Number of Steps Taken")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
Data3[Data3$`Data$steps` == max(Data3$`Data$steps`), ]
```

Due to biases in calculation, we should look for N/A values in out data set, so let's start fresh again and see the summary of steps for the number of N/A rows

```{r}
Data <- read.csv(file.path(path), header=T, sep=",", stringsAsFactors=FALSE)
summary(Data$steps)
```

Let's calculate the number of N/A values

```{r}
sum(is.na(Data))
```

The following for loop loofs for i in 1 to the length of the data, and for each row it checks if it is an N/A value on steps, if it is then it replaces the steps data with the mean of that interval, taken from the previously aggregated mean data.

```{r}
for (i in 1:nrow(Data)){
        if(is.na(Data$steps[i])){
                Data$steps[i] <- Data3[which(Data$interval[i]==Data3$`Data$interval`),]$`Data$steps`
        }
}

sum(is.na(Data))
```

Next we plot the new data with the N/A's filled in:

```{r}
Data$date <- as.Date(Data$date)
g <- ggplot(Data, aes(Data$date, Data$steps))
g + geom_bar(stat="identity", colour = "steelblue", fill = "steelblue", width = 0.7) +
        labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")
```

Then from this new data we take the new mean and median from a newly made aggregate.

```{r}
NewData <- aggregate(Data$steps, list(Data$date), FUN="sum")

mean(NewData$x)
median(NewData$x)
```

The above values are from the newly filled N/A data, what if we compare this new mean/median to the old data's mean/median. As you can see from the values below, the means for the data are the same, but the median for the new data has been shifted up by 1.19 steps.

```{r}
mean(NewData$x)-mean(Data2$`Data$steps`)
median(NewData$x)-median(Data2$`Data$steps`)
```

The code below turns date into a date format, then assigns the days to day names (Monday, Tuesday etc), it then assigns it to a column in our data frame and applies a levels command to change the levels to assign the days to a weekday or a weekend day.

```{r}
Data$date <- as.Date(Data$date)
day <- weekdays(Data$date)
Data$day <- day
levels(Data$day) <- list(weekday=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), weekend=c("Saturday", "Sunday"))
```

Now we can assign the names of the weekend/weekdays through a really inefficient method and use a lattice time series to plot it.

```{r}

FinalHist <- aggregate(Data$steps,list(Data$day,Data$interval), FUN="mean")

FinalHist$Group.1[FinalHist$Group.1=="Monday"] <- "weekday"
FinalHist$Group.1[FinalHist$Group.1=="Tuesday"] <- "weekday"
FinalHist$Group.1[FinalHist$Group.1=="Wednesday"] <- "weekday"
FinalHist$Group.1[FinalHist$Group.1=="Thursday"] <- "weekday"
FinalHist$Group.1[FinalHist$Group.1=="Friday"] <- "weekday"

FinalHist$Group.1[FinalHist$Group.1=="Saturday"] <- "weekend"
FinalHist$Group.1[FinalHist$Group.1=="Sunday"] <- "weekend"

library(lattice)
xyplot(FinalHist$x ~ FinalHist$Group.2 | FinalHist$Group.1, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps")
```

